Using Scipion for Stream Image Processing at Cryo-EM Facilities

### 2018-05-07 0.abstract.tex
final version

Three dimensional electron microscopy is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed. To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}. We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired. This approach helps to detect problems at early stages, saves computing time and provides users with a detailed evaluation of the data quality before the acquisition is finished. Different EM facilities have 
implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests. The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc). This summary may be generated in HTML format and  published through a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities throughout the world and, it is being considered in many more.

>>>>

Three dimensional electron microscopy is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed. To manage such large-scale projects, we had previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}. We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired. This approach helps to detect problems at early stages, saves computing time and provides users with a detailed evaluation of the data quality before the acquisition is finished. Different EM facilities have 
implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests. The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or with a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going on and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc). This summary may be generated in HTML format and  published through a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in seven Cryo-EM facilities throughout the world, and it is being considered in many more.

### 2018-05-07 3.workflow_cnb.tex
final version

Data sets are recorded using \epu. The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS). This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion. \scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks respectively (we will refer to this machine as \scipionbox). \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs. Finally, a second NAS with four 1Gb NICs is linked to the private, LAN and WAN networks (we will refer to this second NAS as output-NAS).

The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copy -or update- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).

@@

\begin{enumerate}
 \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this file and use them to create new predetermined workflows.
 \item  Automatic creation of pdf user's report that contains microscope acquisition parameters and histograms summarizing data resolution and defocus. 
 \item The above mentioned information is stored in a database for further analysis. By default, the system plots the  average resolution and astigmatism of the different acquisitions versus time.
 \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion.
 \item Automatic backup of the \scipion project.
 \item \scipion is able to create live HTML reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information.
\end{enumerate}

>>>>

Data sets are recorded using \epu. The microscope control system has a 70 TB NAS (Network-attached storage) unit for saving movies as they are obtained (we will refer to this NAS as microscope-NAS). This storage unit is shared through a Samba share and connected to the private network using a 10Gb connexion. \scipion is executed on a Linux server which has two 1Gb NICs (network interface controllers) that connect it to the private and the LAN networks respectively (we will refer to this machine as \scipionbox). \scipionbox has 32 CPUs at 2.40GHz, 62 Gb of RAM memory and 2 \textit{Quadro M4000} GPUs. Finally, a second NAS with four 1Gb NICs is linked to the private LAN and WAN networks (we will refer to this second NAS as output-NAS).

The data flow is as follows: (a) the microscope system stores movies in the microscope-NAS, (b) when executing alignment-movie protocols \scipionbox reads the movies from the microscope-NAS but does not store them locally, (c) each hour, output-NAS connects to \scipionbox and copies -or updates- the last \scipion project, (d) at the same time output-NAS may be either executing a script to incrementally backup \scipion project and the movies to a USB disk or serve this information through the Internet and, (e) in parallel, \scipionbox transfers to  output-NAS different HTML reports that may be checked by users or staff with a HTML browser (see \url{http://nolan.cnb.csic.es/scipionbox/lastHTMLReport/}).

@@

\begin{enumerate}
 \item Transparent addition of new workflows. \scipion is able to export any executed workflow as a json file. \emadmin can import this file and use it to create new predetermined workflows.
 \item  Automatic creation of pdf user's report that contains microscope acquisition parameters and histograms summarizing data resolution and defocus. 
 \item The above mentioned information is stored in a database for further analysis. By default, the system plots the  average resolution and astigmatism of the different acquisitions versus time.
 \item Remote management. \emadmin is a client server application, the client maybe executed in a machine different from the one that executes \scipion.
 \item Automatic backup of the \scipion project.
 \item \scipion is able to create live HTML reports on the image acquisition process. \emadmin keeps track of all of them and facilitates the access to this information.
\end{enumerate}

### 2018-05-07 3.workflow_scilifelab.tex
final version

At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion streaming processing. Although there already were using home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf), the microscope operators recognized the importance of a more general
framework. In particular, they were interested in the possibility of modifying easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection.
Although initially the same setup script used at
CNB was adopted, a new one was developed to fully satisfy the facility
requirements. In the following sections we will briefly describe the computational infrastructure at the \scilifelab node and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion. 
%[** thalos falcon III, titan k2 and falcon III]

>>>>

At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion streaming processing. Although the microscope operators were using home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf), they recognized the importance of a more general
framework. In particular, they were interested in the possibility of modifying easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection.
Although initially the same setup script used at
CNB was adopted, a new one was developed to fully satisfy the facility
requirements. In the following sections we will briefly describe the computational infrastructure at the \scilifelab node and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion. 
%[** thalos falcon III, titan k2 and falcon III]

### 2018-05-07 4.future.tex
final version

As part of the EOSCPilot project (\url{https://eoscpilot.eu/cryoem}), we are also improving  how \scipion exports workflows in order to ease reporting of the work done. The goal is to write a workflow file that fully describes the image processing steps, so that if the same input data is provided to the workflow the same results should be obtained. In this way we want to make  the data produced by \scipion more compliant with FAIR guidelines given by the \citet{eu2016:fair}. This workflow file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases like EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}. To facilitate the visualization of the workflow file in any of these databases, we are  developing a  webcomponent (\scipion \emph{workflow viewer}) that will easily allow these repositories to visualize the workflow on their web pages (\url{https://github.com/I2PC/web-workflow-viewer}). Finally, we are implementing a workflow repository. This repository will contain ready-to-use workflows that support a range of use cases.

>>>>

As part of the EOSCPilot project (\url{https://eoscpilot.eu/cryoem}), we are also improving  how \scipion exports workflows in order to ease reporting of the work done. The goal is to write a workflow file that fully describes the image processing steps, so that if the same input data is provided to the workflow the same results should be obtained. In this way we want to make  the data produced by \scipion more compliant with FAIR guidelines given by the \citet{eu2016:fair}. This workflow file could go with the raw data acquired by CryoEM facilities and deposited at common EM databases such as EMPIAR or EMDB \citep{Patwardhan2016:databasesEM}. To facilitate the visualization of the workflow file in any of these databases, we are  developing a  webcomponent (\scipion \emph{workflow viewer}) that will easily allow these repositories to visualize the workflow on their web pages (\url{https://github.com/I2PC/web-workflow-viewer}). Finally, we are implementing a workflow repository. This repository will contain ready-to-use workflows that support a range of use cases.
